\acknowledgements{
I can't say enough to thank my advisor, Mart\'in Farach-Colton. Under his
guidance, I have found countless cool things to learn, awesome problems to work
on, and great people to work with. More than that, him and his family have been
great friends to me, have helped me when I was struggling and helped me grow
into a better person. Most of all, these five years have been so much fun.

I have had the pleasure of working with a truly amazing group of collaborators.
It has been invaluable to discuss systems with Don Porter and theory with
Michael Bender. I have loved working with Ainesh Bakshi, Yizheng Jiao, William
Jannen, Nirjhar Mukherjee, Prashant Pandey, Guido Tagliavini Ponce, Meng-Tsung
Tsai, Jun Yuan and Yang Zhan. Working together with everyone has been one of
my favorite things in graduate school, and is one of the reasons I've been
successful and stuck it through.

I would especially like to thank Rob Johnson, with whom I interned at VMware
Research Group for 2 years, and with whose help I have secured the next phase
of my research career. I am greatly looking forward to working there together
in the future.

I would like to thank everyone else whom I've had the pleasure of working with
at VMware, specifically Vijay Chidambaram, Kapil Chowksey, Yoni Fogel, Abhishek
Gupta, Srinath Premachandran and Amy Tai. Also my time there would not have
been the same without my good friends, Reto Achermann, Amogh Akshintala and
Soujanya Ponnapalli.

I would also like to thank my committee members for their time and support.

I would like to thank my family for their support. My mother has always pushed
me to work hard and live up to my potential, and especially to pursue hard and
interesting research. My brother, Oliver, has spent many hours chatting with me
about different problems. And I would also like to thank my mother-in-law,
Zsuzsanna, for encouraging and supporting me (not to mention letting me crash
with her near Rutgers!).

Finally, I would like to thank my partner, Hanna, who has been there for me for
the last 15 years. She has put up with me when I've struggled and has always
been there to stand me up on my feet again. I know I could not have done this
without her, and I hope that I can support her on her upcoming academic journey
as well as she has on mine.
}

