\section{Related Work}\label{sec:fsa-related}
Prior work on file system aging falls into three categories:
techniques for artificially inducing aging, for
measuring aging, and for mitigating aging.

\subsection{Creating Aged File Systems}

The seminal work of Smith and Seltzer~\cite{DBLP:conf/sigmetrics/SmithS97}
created a methodology for simulating and measuring aging on a file
system---leading to more representative benchmark results than running on a
new, empty file system.  The study is based on data  collected from daily
snapshots of over fifty real file systems from five servers over durations
ranging from one to three years.  An overarching goal of Smith and Seltzer's
work was to evaluate file systems with representative levels of aging.

Other tools have been subsequently developed for synthetically aging a file
system.  In order to measure NFS performance,
TBBT~\cite{DBLP:conf/fast/ZhuCC05} was designed to synthetically age a disk to
create a initial state for NFS trace replay.

The Impressions framework~\cite{DBLP:journals/tos/AgrawalAA09} was designed so
that users can synthetically age a file system by setting a small number of
parameters, such as the organization of the directory hierarchy.  Impressions
also lets users specify a target layout score for the resulting image.

Both TBBT and Impressions create file systems with a specific level of
fragmentation, whereas our study identifies realistic workloads that induce
fragmentation. 

\subsection{Measuring Aged File Systems}

Smith and Seltzer also introduced a \defn{layout score} for studying aging,
which was used by subsequent
studies~\cite{DBLP:conf/mascots/AhnKCP02,DBLP:journals/tos/AgrawalAA09}. Their
layout score is the fraction of file blocks that are placed in consecutive
physical locations on the disk. We introduce a variation of this measure, the
\defn{dynamic layout score} in \cref{sec:fsa-metrics}.

The \defn{degree of fragmentation} (\defn{DoF}) is used in the study of
fragmentation in mobile devices~\cite{DBLP:conf/hotstorage/JiCSWLX16}. DoF is
the ratio of the actual number of extents, or ranges of contiguous physical
blocks, to the ideal number of extents.  Both the layout score and DoF measure
how one file is fragmented.

Several studies have reported file system statistics such as number of files,
distributions of file sizes and types, and organization of file system
namespaces~\cite{DBLP:journals/tos/AgrawalBDL07,DBLP:conf/mascots/Downey01,DBLP:conf/usenix/RoselliLA00}.
These statistics can inform parameter choices in aging frameworks like TBBT and
Impressions~\cite{DBLP:conf/fast/ZhuCC05,DBLP:journals/tos/AgrawalAA09}.
